---
title: "Chapter3.5"
author: "Min-Yao"
date: "2021/1/16"
output: 
  html_document: 
    keep_md: yes
---

## 3.5. CLASSIFYING NEWSWIRES: A MULTICLASS CLASSIFICATION EXAMPLE

### 3.5.1. The Reuters dataset

#### Listing 3.10. Loading the Reuters dataset

```{r}
library(keras)

reuters <- dataset_reuters(num_words = 10000)
c(c(train_data, train_labels), c(test_data, test_labels)) %<-% reuters
```

```{r}
length(train_data)
```

```{r}
length(test_data)
```

```{r}
train_data[[1]]
```

#### Listing 3.11. Decoding newswires back to text

```{r}
word_index <- dataset_reuters_word_index()
reverse_word_index <- names(word_index)
names(reverse_word_index) <- word_index
decoded_newswire <- sapply(train_data[[1]], function(index) {
  word <- if (index >= 3) reverse_word_index[[as.character(index - 3)]]    
  if (!is.null(word)) word else "?"
})
```

```{r}
train_labels[[1]]
```

### 3.5.2. Preparing the data

#### Listing 3.12. Encoding the data

```{r}
vectorize_sequences <- function(sequences, dimension = 10000) {
  results <- matrix(0, nrow = length(sequences), ncol = dimension)
  for (i in 1:length(sequences))
    results[i, sequences[[i]]] <- 1
  results
}

x_train <- vectorize_sequences(train_data)          
x_test <- vectorize_sequences(test_data)
```

```{r}
to_one_hot <- function(labels, dimension = 46) {
  results <- matrix(0, nrow = length(labels), ncol = dimension)
  for (i in 1:length(labels))
    results[i, labels[[i]] + 1] <- 1
  results
}

one_hot_train_labels <- to_one_hot(train_labels)
one_hot_test_labels <- to_one_hot(test_labels)
```

```{r}
one_hot_train_labels <- to_categorical(train_labels)
one_hot_test_labels <- to_categorical(test_labels)
```

### 3.5.3. Building your network

#### Listing 3.13. Model definition

```{r}
model <- keras_model_sequential() %>%
  layer_dense(units = 64, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dense(units = 46, activation = "softmax")
```

#### Listing 3.14. Compiling the model

```{r}
model %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("accuracy")
)
```

### 3.5.4. Validating your approach

#### Listing 3.15. Setting aside a validation set

```{r}
val_indices <- 1:1000

x_val <- x_train[val_indices,]
partial_x_train <- x_train[-val_indices,]

y_val <- one_hot_train_labels[val_indices,]
partial_y_train = one_hot_train_labels[-val_indices,]
```

#### Listing 3.16. Training the model

```{r}
history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)
```

```{r}
plot(history)
```

#### Listing 3.18. Retraining a model from scratch

```{r}
model <- keras_model_sequential() %>%
  layer_dense(units = 64, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dense(units = 46, activation = "softmax")

model %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("accuracy")
)

history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 9,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

results <- model %>% evaluate(x_test, one_hot_test_labels)
```

```{r}
results
```

```{r}
test_labels_copy <- test_labels
test_labels_copy <- sample(test_labels_copy)
length(which(test_labels == test_labels_copy)) / length(test_labels)
```

### 3.5.5. Generating predictions on new data

#### Listing 3.19. Generating predictions for new data

```{r}
predictions <- model %>% predict(x_test)
```

```{r}
dim(predictions)
```

```{r}
sum(predictions[1,])
```

```{r}
which.max(predictions[1,])
```

### 3.5.6. A different way to handle the labels and the loss

```{r}
model %>% compile(
  optimizer = "rmsprop",
  loss = "sparse_categorical_crossentropy",
  metrics = c("accuracy")
)
```

### 3.5.7. The importance of having sufficiently large intermediate layers

#### Listing 3.20. A model with an information bottleneck

```{r}
model <- keras_model_sequential() %>%
  layer_dense(units = 64, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 4, activation = "relu") %>%
  layer_dense(units = 46, activation = "softmax")

model %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("accuracy")
)

model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 128,
  validation_data = list(x_val, y_val)
)
```
### 3.5.8. Further experiments

1. Try using larger or smaller layers: 32 units, 128 units, and so on.


2. You used two hidden layers. Now try using a single hidden layer, or three hidden layers.


### 3.5.9. Wrapping up

## 3.6. PREDICTING HOUSE PRICES: A REGRESSION EXAMPLE

Don’t confuse regression with the algorithm logistic regression. Confusingly, logistic regression isn’t a regression algorithm—it’s a classification algorithm.

### 3.6.1. The Boston Housing Price dataset

#### Listing 3.21. Loading the Boston housing dataset

```{r}
library(keras)

dataset <- dataset_boston_housing()
c(c(train_data, train_targets), c(test_data, test_targets)) %<-% dataset
```

```{r}
str(train_data)
```

```{r}
str(test_data)
```

```{r}
str(train_targets)
```

### 3.6.2. Preparing the data

#### Listing 3.22. Normalizing the data

```{r}
mean <- apply(train_data, 2, mean)                
std <- apply(train_data, 2, sd)
train_data <- scale(train_data, center = mean, scale = std)
test_data <- scale(test_data, center = mean, scale = std)
```

### 3.6.3. Building your network

#### Listing 3.23. Model definition

```{r}
build_model <- function() {                               
  model <- keras_model_sequential() %>%
    layer_dense(units = 64, activation = "relu",
                input_shape = dim(train_data)[[2]]) %>%
    layer_dense(units = 64, activation = "relu") %>%
    layer_dense(units = 1)
  model %>% compile(
    optimizer = "rmsprop",
    loss = "mse",
    metrics = c("mae")
  )
}
```

### 3.6.4. Validating your approach using K-fold validation

#### Figure 3.9. 3-fold cross-validation

#### Listing 3.24. K-fold validation

```{r}
k <- 4
indices <- sample(1:nrow(train_data))
folds <- cut(indices, breaks = k, labels = FALSE)

num_epochs <- 100
all_scores <- c()
for (i in 1:k) {
  cat("processing fold #", i, "\n")

  val_indices <- which(folds == i, arr.ind = TRUE)                     #1
  val_data <- train_data[val_indices,]
  val_targets <- train_targets[val_indices]
  partial_train_data <- train_data[-val_indices,]                      #2
  partial_train_targets <- train_targets[-val_indices]

  model <- build_model()                                               #3

  model %>% fit(partial_train_data, partial_train_targets,             #4
                epochs = num_epochs, batch_size = 1, verbose = 0)

  results <- model %>% evaluate(val_data, val_targets, verbose = 0)    #5
  all_scores <- c(all_scores, results$mean_absolute_error)
}
```


```{r}
all_scores
```

```{r}
mean(all_scores)
```

