---
title: "Chapter 3"
author: "Min-Yao Jhu"
date: "1/16/2021"
output: 
  html_document: 
    keep_md: yes
---

## load lib

```{r}
library(keras)
library(reticulate)
use_condaenv("r-reticulate")
```
# Chapter 3. Getting started with neural networks

## 3.1. ANATOMY OF A NEURAL NETWORK

### 3.1.1. Layers: the building blocks of deep learning

```{r}
layer <- layer_dense(units = 32, input_shape = c(784))
```

```{r}
model <- keras_model_sequential() %>%
  layer_dense(units = 32, input_shape = c(784)) %>%
  layer_dense(units = 32)
```

### 3.1.2. Models: networks of layers

### 3.1.3. Loss functions and optimizers: keys to configuring the learning process

## 3.2. INTRODUCTION TO KERAS

### 3.2.1. Keras, TensorFlow, Theano, and CNTK

### 3.2.2. Installing Keras

### 3.2.3. Developing with Keras: a quick overview

## 3.3. SETTING UP A DEEP-LEARNING WORKSTATION

### 3.3.1. Getting Keras running: two options

### 3.3.2. Running deep-learning jobs in the cloud: pros and cons

### 3.3.3. What is the best GPU for deep learning?

## 3.4. CLASSIFYING MOVIE REVIEWS: A BINARY CLASSIFICATION EXAMPLE

### 3.4.1. The IMDB dataset

#### Listing 3.1. Loading the IMDB dataset

```{r}
library(keras)

imdb <- dataset_imdb(num_words = 10000)
c(c(train_data, train_labels), c(test_data, test_labels)) %<-% imdb
```

```{r}
str(train_data[[1]])
```

```{r}
train_labels[[1]]
```

```{r}
max(sapply(train_data, max))
```

### 3.4.2. Preparing the data

#### Listing 3.2. Encoding the integer sequences into a binary matrix

```{r}
vectorize_sequences <- function(sequences, dimension = 10000) {
  results <- matrix(0, nrow = length(sequences), ncol = dimension)    
  for (i in 1:length(sequences))
    results[i, sequences[[i]]] <- 1                                   
  results
}

x_train <- vectorize_sequences(train_data)
x_test <- vectorize_sequences(test_data)
```

```{r}
str(x_train[1,])
```

```{r}
y_train <- as.numeric(train_labels)
y_test <- as.numeric(test_labels)
```

### 3.4.3. Building your network

```{r}
library(keras)

model <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")
```

#### Listing 3.4. Compiling the model

```{r}
model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)
```

#### Listing 3.5. Configuring the optimizer

```{r}
model %>% compile(
  optimizer = optimizer_rmsprop(lr=0.001),
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)
```

#### Listing 3.6. Using custom losses and metrics

```{r}
model %>% compile(
  optimizer = optimizer_rmsprop(lr = 0.001),
  loss = loss_binary_crossentropy,
  metrics = metric_binary_accuracy
)
```

### 3.4.4. Validating your approach

#### Listing 3.7. Setting aside a validation set

```{r}
val_indices <- 1:10000

x_val <- x_train[val_indices,]
partial_x_train <- x_train[-val_indices,]
y_val <- y_train[val_indices]
partial_y_train <- y_train[-val_indices]
```

#### Listing 3.8. Training your model

```{r}
model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)
```

```{r}
str(history)
```

```{r}
plot(history)
```
#### Training history with the plot() method

```{r}
history_df <- as.data.frame(history)
str(history_df)
```
#### Listing 3.9. Retraining a model from scratch

```{r}
model <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

model %>% fit(x_train, y_train, epochs = 4, batch_size = 512)
results <- model %>% evaluate(x_test, y_test)
```

```{r}
results
```

### 3.4.5. Using a trained network to generate predictions on new data

### 3.4.6. Further experiments

1. You used two hidden layers. Try using one or three hidden layers, and see how doing so affects validation and test accuracy.

> 1 layer, a little bit better

```{r}
model1 <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 1, activation = "sigmoid")

model1 %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

model1 %>% fit(x_train, y_train, epochs = 4, batch_size = 512)
results <- model1 %>% evaluate(x_test, y_test)
results
```

> 3 layers, not better

```{r}
model3 <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

model3 %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

model3 %>% fit(x_train, y_train, epochs = 4, batch_size = 512)
results <- model3 %>% evaluate(x_test, y_test)
results
```

2. Try using layers with more hidden units or fewer hidden units: 32 units, 64 units, and so on.

> 32 units, not better, about the same

```{r}
model_32 <- keras_model_sequential() %>%
  layer_dense(units = 32, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

model_32 %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

model_32 %>% fit(x_train, y_train, epochs = 4, batch_size = 512)
results <- model_32 %>% evaluate(x_test, y_test)
results
```

> 64 units, worse

```{r}
model_64 <- keras_model_sequential() %>%
  layer_dense(units = 64, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

model_64 %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

model_64 %>% fit(x_train, y_train, epochs = 4, batch_size = 512)
results <- model_64 %>% evaluate(x_test, y_test)
results
```



3. Try using the mse loss function instead of binary_crossentropy.

> mse loss function, not better

```{r}
model_MSE <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

model_MSE %>% compile(
  optimizer = "rmsprop",
  loss = "mse",
  metrics = c("accuracy")
)

model_MSE %>% fit(x_train, y_train, epochs = 4, batch_size = 512)
results <- model_MSE %>% evaluate(x_test, y_test)
results
```


4. Try using the tanh activation (an activation that was popular in the early days of neural networks) instead of relu.

> tanh activation, not better

```{r}
model_tanh <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "tanh", input_shape = c(10000)) %>%
  layer_dense(units = 16, activation = "tanh") %>%
  layer_dense(units = 1, activation = "sigmoid")

model_tanh %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

model_tanh %>% fit(x_train, y_train, epochs = 4, batch_size = 512)
results <- model_tanh %>% evaluate(x_test, y_test)
results
```




### 3.4.7. Wrapping up

