---
title: "OJ"
author: "Min-Yao"
date: "2021/1/16"
output: 
  html_document: 
    keep_md: yes
---

## OJ

```{r}
library(tidyverse)
library(ISLR)
library(tree)
data(OJ)
#?OJ
```

```{r}
head(OJ)
summary(OJ)
```

```{r}
OJ %>% select(StoreID, STORE) %>% head()
```
> StoreID and STORE are redundant
> because store is categorical we need to turn that into a series of dummy variables.

```{r}
store_cat <- OJ %>% select(StoreID) %>%
  mutate(row=1:nrow(.),data=1) %>%
  pivot_wider(id_cols = row, names_from=StoreID, values_from=data, values_fill=0, names_prefix="Store")
head(store_cat)
```

```{r}
OJ <- OJ %>% select(-StoreID, -STORE, -Store7) %>% cbind(store_cat)
head(OJ)
```

## (a) Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.

```{r}
set.seed(10)
train <- sample(1:nrow(OJ), size = 800)
OJ.train <- OJ[train,]
OJ.test <- OJ[-train,]
str(OJ.test)
str(OJ.train)
```

## (b) Fit a tree to the training data, with Purchase as the response and the other variables as predictors. Use the summary() function to produce summary statistics about the tree, and describe the results obtained. What is the training error rate? How many terminal nodes does the tree have?

```{r}
tree.OJ=tree(Purchase~.,OJ,subset=train)
summary(tree.OJ)
```
> Number of terminal nodes:  7

> Misclassification error rate: 0.1775

## (e) Predict the response on the test data, and produce a confusion matrix comparing the test labels to the predicted test labels. What is the test error rate?

```{r}
tree.pred=predict(tree.OJ,OJ.test,type="class")
purchase.test=OJ.test$Purchase
table(tree.pred,purchase.test)
(135+88)/270
1-((135+88)/270)
```

> Accuracy is 82.59%

## Deep learning method

```{r}
library(keras)
library(reticulate)
library(tensorflow)
use_condaenv("r-reticulate")
```

> prepare label

```{r}
OJ.train.label <- OJ.train %>% select(Purchase) %>% mutate(Purchase=ifelse(Purchase=="CH", 0, 1)) %>% pull(Purchase) # pull take it out as a vector, the same as oj[["purchase"]]
str(OJ.train.label)
OJ.test.label <- OJ.test %>% select(Purchase) %>% mutate(Purchase=ifelse(Purchase=="CH", 0, 1)) %>% pull(Purchase)
str(OJ.test.label)
```

> prepare data

```{r}
OJ.train <- OJ.train %>% select(-Purchase)
str(OJ.train)
OJ.test <- OJ.test %>% select(-Purchase)
str(OJ.test)
```

> scale data

```{r}
oj.mean <- apply(OJ.train, 2, mean)
oj.std <- apply(OJ.train, 2, sd)

OJ.train <- scale(OJ.train, center=oj.mean, scale=oj.std)
OJ.test <- scale(OJ.test, center=oj.mean, scale=oj.std)

str(OJ.train)
str(OJ.test)
```

> Setting aside a validation set

```{r}
set.seed(10)
val_indices <- sample(1:nrow(OJ.train), size = 100)

OJ.train_val <- OJ.train[val_indices,]
partial_OJ.train <- OJ.train[-val_indices,]

# class(OJ.train) # to know the 

# OJ.test <- data.matrix(OJ.test, rownames.force = NA)
# OJ.train <- data.matrix(OJ.train, rownames.force = NA)
# OJ.train_val <- data.matrix(OJ.train_val, rownames.force = NA)
# partial_OJ.train <- data.matrix(partial_OJ.train, rownames.force = NA)

# matrix and dataframe are different
# matrix need to be the same element (as )

OJ.train_val.label <- OJ.train.label[val_indices]
partial_OJ.train.label <- OJ.train.label[-val_indices]

dim(OJ.train_val)
dim(partial_OJ.train)

class(OJ.train_val)
class(partial_OJ.train)
class(OJ.train_val.label)
class(partial_OJ.train.label)

str(OJ.train_val)
str(partial_OJ.train)
str(OJ.train_val.label)
str(partial_OJ.train.label)

```

## Try different hyperparamaters

### variant 1: larger or smaller layers

```{r}
define_model <- function(powerto) {
  network_int <- keras_model_sequential() %>% 
    layer_dense(units = 2^powerto, activation = "relu", input_shape = ncol(OJ.train)) %>% 
    layer_dense(units = 2^powerto, activation = "relu") %>% 
    layer_dense(units = 1, activation = "sigmoid")
  
  network_int %>% compile(
    optimizer = "rmsprop",
    loss = "binary_crossentropy",
    metrics = c("accuracy")
  )
}
```

```{r}
run_model <- function(network, epochs) {
  network %>% fit(
    partial_OJ.train,
    partial_OJ.train.label,
    epochs = epochs,
    batch_size = 256,
    validation_data = list(OJ.train_val, OJ.train_val.label)
  )
}
```

> units = 8

```{r}
define_model(3) %>%
  run_model(100) -> history_3

history_3 %>%
  plot()
```

> units = 16

```{r}
define_model(4) %>%
  run_model(100) -> history_4

history_4 %>%
  plot()
```

> units = 32

```{r}
define_model(5) %>%
  run_model(100) -> history_5

history_5 %>%
  plot()
```
> units = 64

```{r}
define_model(6) %>%
  run_model(100) -> history_6

history_6 %>%
  plot()
```

### variant 2: how many layers

```{r}
define_model <- function(nlayers, powerto) {
  
  # input layer
  network <- keras_model_sequential() %>% 
    layer_dense(units = 2^powerto, activation = "relu", input_shape = ncol(OJ.train)) 
  
  # additional layers
  if (nlayers>1) {
  map(2:nlayers, ~ network %>% 
        layer_dense(units = 2^powerto, activation = "relu")
  )
  }
  
  # output layer
  network %>% 
    layer_dense(units = 1, activation = "sigmoid")
  
  # compile it
  network %>% compile(
    optimizer = "rmsprop",
    loss = "binary_crossentropy",
    metrics = c("accuracy")
  )
  
}
```

> 1 layer, units = 16

```{r}
define_model(1,4) %>%
  run_model(100) -> history_1_4

history_1_4 %>%
  plot()
```

> 2 layer, units = 16

```{r}
define_model(2,4) %>%
  run_model(100) -> history_2_4

history_2_4 %>%
  plot()
```

> 3 layer, units = 16

```{r}
define_model(3,4) %>%
  run_model(100) -> history_3_4

history_3_4 %>%
  plot()
```

### I decide to use units = 16 and 2 layers to run the model.

```{r}
model <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = ncol(OJ.train) ) %>% 
  layer_dense(units = 16, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")


model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

system.time(history <- model %>% fit(
  partial_OJ.train,
  partial_OJ.train.label,
  epochs = 100,
  batch_size = 256,
  validation_data = list(OJ.train_val, OJ.train_val.label)
))
```

```{r}
str(history)
```

```{r}
plot(history)
```

> I choose 56, run with 56 epochs and full training:

```{r}
model <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = ncol(OJ.train) ) %>% 
  layer_dense(units = 16, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")


model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

system.time(history <- model %>% fit(
  OJ.train,
  OJ.train.label,
  epochs = 56,
  batch_size = 256
))
```


```{r}
plot(history)
```


```{r}
results <- model %>% evaluate(OJ.test, OJ.test.label)

results
```

> Accuracy is 84.81%, which is a little bit better than the trees (82.59%).

# batch size also power of 2
# smaller train better, but also slower
